RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-01-09 paper/RDF validation requirements/RDF validation requirements.tex
RDF citations

In this paper, we describe each requirement within the RDF validation requirements database in detail (sections 3-76).
Additional descriptions can be found directly in the database.
Each requirement corresponds to an RDF constraint which may be expressible by multiple constraint languages.
For each requirement, we represent some examples in different constraint languages.

@@

\ms{String Operations} corresponds to the requirement
\ms{R-194-PROVIDE-STRING-} \ms{FUNCTIONS-FOR-RDF-LITERALS}.
Some constraints require building new strings out of other strings.
Calculating the string length would also be another constraint of this type.

@@

\ms{Context-Specific Valid Classes} corresponds to the requirements \\
\ms{R-209-VALID-CLASSES}.
What types of resources (rdf:type) are valid in a specific context? 
Context can be an input stream, a data creation function, or an API.

@@

\ms{Context-Specific Valid Properties} corresponds to the requirement
\ms{R-210-} \ms{VALID-PROPERTIES}.
What properties can be used within this context? 
Context can be an data receipt function, data creation function, or API.

@@

\ms{Default Values} corresponds to the requirements
\ms{R-31-DEFAULT-VALUES-OF-RDF-} \ms{OBJECTS} and
\ms{R-38-DEFAULT-VALUES-OF-RDF-LITERALS}.
Default values for objects and literals are inferred automatically.
It should be possible to declare the default value for a given property, e.g. so that input forms can be pre-populated and to insert a required property that is missing in a web service call.

@@

\begin{ex}
# rule (SPIN)
# -----------
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {            
                    ?this laserSwordColor "blue"^^xsd:string ;
                    ?this numberLaserSwords "1"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a Jedi .            
                } """ ; ] .
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {
                    ?this laserSwordColor "red"^^xsd:string ;
                    ?this numberLaserSwords "2"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a Sith .            
                } """ ; ] .
\end{ex}

\begin{ex}
# data:
Joda a Jedi .
DarthSidious a Sith .
\end{ex}

\begin{ex}
# inferred triples:
Joda 
    laserSwordColor "blue"^^xsd:string ;
    numberLaserSwords "1"^^xsd:nonNegativeInteger .
DarthSidious 
    laserSwordColor "red"^^xsd:string ;
    numberLaserSwords "2"^^xsd:nonNegativeInteger .
\end{ex}

@@

\ms{Mathematical Operations} corresponds to the requirements
\begin{itemize}
	\item \ms{R-42-MATHEMATICAL-OPERATIONS} and
  \item \ms{R-41-STATISTICAL-COMPUTATIONS}.
\end{itemize}
Examples are:
\begin{itemize}
	\item adding 2 dates
	\item add number of days to start date
	\item area = width * height
	\item Statistical Computations: average, mean, sum
\end{itemize}

@@

\ms{Language Tag Matching} corresponds to the requirement \\
\ms{R-47-LANGUAGE-TAG-MATCHING}.

@@

\ms{Language Tag Cardinality} corresponds to the requirements
\begin{itemize}
	\item \ms{R-49-RDF-LITERALS-HAVING-AT-MOST-ONE-LANGUAGE-TAG} and
	\item \ms{R-48-MISSING-LANGUAGE-TAGS}.
\end{itemize}

@@

Check that no language is used more than once per property

\begin{ex}
# DQTP:
SELECT DISTINCT ?s WHERE { ?s %%P1%% ?c
    BIND ( lang(?c) AS ?l )
    FILTER (isLiteral (?c) && lang(?c) = %%V1%%)}
GROUP BY ?s HAVING COUNT (?l) > 1
\end{ex}

This corresponds to the test pattern \ms{ONELANGPattern} \cite{Kontokostas2014}.
A literal value should contain at most 1 literal for a language. 
\ms{P1} is the property containing the literal and \ms{V1} is the language we want to check.

\begin{ex}
# test binding (DQTP):
P1 => foaf:name
V1 => en
\end{ex}

A single English (“en”) \ms{foaf:name}.

\begin{ex}
# valid data:
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en .
\end{ex}

\begin{ex}
# invalid data:
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en ;
    foaf:name 'Leia'@en .
\end{ex}

@@

\end{document}

>>>>

In this paper, we describe each requirement within the RDF validation requirements database in detail (sections 3-76).
Each requirement corresponds to an RDF constraint which may be expressible by multiple constraint languages.
For each requirement, we represent some examples in different constraint languages.

@@

\begin{itemize}
	\item R-194-PROVIDE-STRING-FUNCTIONS-FOR-RDF-LITERALS
\end{itemize}

\begin{itemize}
	\item Some constraints require building new strings out of other strings
\end{itemize}

\begin{itemize}
	\item string length
\end{itemize}

@@

\begin{itemize}
	\item R-209-VALID-CLASSES
\end{itemize}

\begin{itemize}
	\item What types of resources (rdf:type) are valid in this context? Context can be an input stream, a data creation function, or an API.
\end{itemize}

@@

\begin{itemize}
	\item R-210-VALID-PROPERTIES
\end{itemize}

\begin{itemize}
	\item What properties can be used within this context? Context can be an data receipt function, data creation function, API, etc.
\end{itemize}

@@

\begin{itemize}
	\item R-31-DEFAULT-VALUES-OF-RDF-OBJECTS
  \item R-38-DEFAULT-VALUES-OF-RDF-LITERALS
\end{itemize}

\begin{itemize}
	\item default values for RDF literals
	\item default values for RDF objects
	\item default values are inferred automatically
	\item It should be possible to declare the default value for a given property, e.g. so that input forms can be pre-populated and to insert a required property that is missing in a web service call. 
\end{itemize}

@@

\begin{ex}
# rule (SPIN)
# -----------
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {            
                    ?this :laserSwordColor "blue"^^xsd:string ;
                    ?this :numberLaserSwords "1"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a :Jedi .            
                } """ ; ] .
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {
                    ?this :laserSwordColor "red"^^xsd:string ;
                    ?this :numberLaserSwords "2"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a :Sith .            
                } """ ; ] .
\end{ex}

\begin{ex}
# data
# ----
:Joda a :Jedi .
:DarthSidious a :Sith .
\end{ex}

\begin{ex}
# inferred triples
# ----------------
:Joda 
    :laserSwordColor "blue"^^xsd:string ;
    :numberLaserSwords "1"^^xsd:nonNegativeInteger .
:DarthSidious 
    :laserSwordColor "red"^^xsd:string ;
    :numberLaserSwords "2"^^xsd:nonNegativeInteger .
\end{ex}

@@

\begin{itemize}
	\item R-42-MATHEMATICAL-OPERATIONS
	\item R-41-STATISTICAL-COMPUTATIONS
\end{itemize}

\begin{itemize}
	\item adding 2 dates
	\item add number of days to start date
	\item area = width * height
	\item Statistical Computations: average, mean, sum
\end{itemize}

@@

\begin{itemize}
	\item R-47-LANGUAGE-TAG-MATCHING
\end{itemize}

@@

\begin{itemize}
	\item R-49-RDF-LITERALS-HAVING-AT-MOST-ONE-LANGUAGE-TAG
	\item R-48-MISSING-LANGUAGE-TAGS
\end{itemize}

@@

check that no language is used more than once per property

\begin{ex}
# DQTP
# ----
SELECT DISTINCT ?s WHERE { ?s %%P1%% ?c
    BIND ( lang(?c) AS ?l )
    FILTER (isLiteral (?c) && lang(?c) = %%V1%%)}
GROUP BY ?s HAVING COUNT (?l) > 1
\end{ex}

\begin{itemize}
	\item ONELANGPattern \cite{Kontokostas2014} 
  \item A literal value should contain at most 1 literal for a language. 
  \item P1 is the property containing the literal and V1 is the language we want to check.
\end{itemize}

\begin{ex}
# test binding (DQTP)
---------------------
P1 => foaf:name
V1 => en
\end{ex}

\begin{itemize}
	\item a single English (“en”) foaf:name
\end{itemize}

\begin{ex}
# valid data
# ----------
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en .
\end{ex}

\begin{ex}
# invalid data
# ------------
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en ;
    foaf:name 'Leia'@en .
\end{ex}

@@
